Work Fragmentation in Developer Interaction Data

### 2016-03-15 paper_2.tex
Complete RQ2

In this section we analyze whether or not the duration of the interruptions is a factor in the relationship between interruptions and developer productivity. Under the hypothesis that prolonged interruptions require more time to recover, we built two groups of sessions based on the proportion of long interruptions. As we described previously we consider that prolonged interruption are those whose duration is $\geq$ 12 minutes, according to previous work \cite{GM04, KaptelininN07}. Prolonged interruptions could be due to an extended interruption, such as a coffee break or a meeting.

@@

Following the conclusions on RQ1, the number of interruptions may affect the results if we ignore that variable. So, we split every group in accordance to the number of interruptions; this time, instead of taking the quantiles of the distribution to create groups we created a group for every number of interruptions up to seven. Since the results in RQ1 show that as the number of interruptions increases the difference between the groups minimizes, we believe that having more distinctive groups instead of groups created by ranges is also valid. Thus, for the three metrics, we created the groups of sessions according to the number on interruptions, between one and seven, en then we split every group in low and high, according to the proportion of prolonged interruptions.

The Figure \ref{fig:dur_mylyn} shows the median of the metrics when grouping the sessions by the number of interruptions and the proportion of prolonged interruptions with Mylyn, and the Figure \ref{fig:dur_udc} with UDC. In this context, it is better to have a low number of interruptions and a low proportion of prolonged interruptions, and the results agree with this statement. First, we observe that (as we conclude in RQ1) when the number of interruption increases the median of the metric decreases until reaching a point (approximately four interruptions) where the change is minimal. Second, almost in every group the median of groups with low proportion of prolonged interruptions is greater than in groups with a high proportion. The latter observation is more clear in the editions per minute metric in both datasets.

%For example, when having between 1 to 3 interruptions and a low proportion the metrics show higher values when compared to having $>$ 5 interruptions and a high proportion of prolonged interruptions. Moreover, if we focus on a category we can see a gradual decrease as the number of interruptions increases, which lead us to conclude that the proportion of prolonged interruptions has an effect on the developer productivity. In the Figures the color of lines represent a classification of the proportion of prolonged interruptions, being the green line the \textit{low} type, blue the \textit{medium} and red the \textit{high}.

%The difference between categories seems to decrease when reaching higher values. There is a clear difference between the low and medium categories of proportions, but the difference between medium and high is minor. The same happens when comparing the categories [2-3] and [4-6] of the number of interruptions, and the categories [4-6] and $>$6. In other words, as the value of the metrics get higher, the negative effect (according to these metrics) changes at a very low rhythm.

%Sessions with only one interruption do not belong to the medium category, for they can only contain either one short interruption (low category) or one prolonged interruption (high category). So, in the Tables the medium category shows an NA in sessions with one interruption. Also, there is an important difference in this instance as the metrics show better results in sessions in the low category, which is another evidence of the detrimental effect of prolonged interruptions.

%Comparing the results between the datasets, UDC seems to show a more clear difference between groups. The sessions with a low proportion of prolonged interruption has a greater difference (higher productivity) than sessions with medium proportion. In Mylyn this difference is more subtle but can be identified, although when having more than 6 interruptions there is not a real difference. We observed that having a high number of prolonged interruptions is more detrimental to productivity, and we can reach the same conclusions with both datasets and with the three metrics.

\begin{figure}[!ht]
\centering
\begin{tabular}{c}
\minipage{0.33\textwidth}
\includegraphics[width=\linewidth,clip=, angle=0]{figures/mylyn_lines_dur_emin}  
\endminipage\hfill
\minipage{0.33\textwidth}
\includegraphics[width=\linewidth,clip=, angle=0]{figures/mylyn_lines_dur_smin}  
\endminipage\hfill
\minipage{0.33\textwidth}
\includegraphics[width=\linewidth,clip=, angle=0]{figures/mylyn_lines_dur_eratio} 
\endminipage\hfill
\end{tabular}
\caption{Boxplots showing the relation between the number of edits and selection events per minute, the edit ratio, and the number of interruptions in Mylyn. }
\label{fig:dur_mylyn}
\end{figure}

\begin{figure}[!ht]
	\centering
	\begin{tabular}{c}
		\minipage{0.33\textwidth}
		\includegraphics[width=\linewidth,clip=, angle=0]{figures/udc_lines_dur_emin}  \\
		\endminipage\hfill
		\minipage{0.33\textwidth}
		\includegraphics[width=\linewidth,clip=, angle=0]{figures/udc_lines_dur_smin}  \\
		\endminipage\hfil
		\minipage{0.33\textwidth}
		\includegraphics[width=\linewidth,clip=, angle=0]{figures/udc_lines_dur_eratio} \\
		\endminipage\hfill
	\end{tabular}
	\caption{Boxplots showing the relation between the number of editions and selections per minute, the edit ratio, and the number of interruptions in UDC. }
	\label{fig:dur_udc}
\end{figure}

The Tables \ref{tbl:effect_size_dur_mylyn} and \ref{tbl:effect_size_dur_udc} show a test performed in every group of interruptions, measuring the effect of sessions with low and high proportion of prolonged interruptions.

\begin{table}[ht!]
	\tiny
	\renewcommand{\arraystretch}{1.3}
	\caption{Effect size and statistical significance of the relationships between edits and selection per minute, edit ratio, and duration of interruptions in Mylyn} %, and editions and selections.}
	\label{tbl:effect_size_dur_mylyn}
	\centering
	\begin{tabular}{l | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm}} 
		Group & 1 & 2 & 3 & 4 & 5 & 6 & 7  \\  
		\hline
		\multicolumn{8}{c}{\textbf{Edits}} \\
		\hline
		U-test (adjusted) & 5.6e-10 & 6.5e-07 & 2.9e-04 & 7.4e-03 & 1.2e-03 & 3.2e-01 &  3.9e-01\\
		\hline
		Cohen's $d$ & 0.62 & 0.38 & 0.20 & 0.08 & 0.21 & 0.17 & 0.45 \\
		\hline

@@

		\multicolumn{8}{c}{\textbf{Edit ratio}} \\
		\hline 
		U-test (adjusted) & 2.6e-06 & 5.7e-05 & 2.3e-02 & 6.4e-01 & 2.3e-02 & 4.9e-01 & 1.5e-01 \\
		\hline 
		Cohen's $d$ & 0.52 & 0.40 & 0.21 & 0.06 & 0.41 & 0.04 & 0.56  \\
		\hline

@@

\begin{table}[ht!]
	\tiny
	\renewcommand{\arraystretch}{1.3}
	\caption{Effect size and statistical significance of the relationships between edits and selection per minute, edit ratio, and duration of interruptions in UDC} %, and editions and selections.}
	\label{tbl:effect_size_dur_udc}
	\centering
	\begin{tabular}{l | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm}} 
		Group & 1 & 2 & 3 & 4 & 5 & 6 & 7  \\  
		\hline
		\multicolumn{8}{c}{\textbf{Edits}} \\
		\hline
		U-test (adjusted) & 1.9e-20 & 4.0e-13 & 1.1e-03 & 2.7e-02 & 7.2e-08 & 5.5e-02 & 6.5e-06\\
		\hline
		Cohen's $d$ & 0.28 & 0.25 & 0.09 & 0.03 & 0.11 & 0.01 & 0.15 \\
		\hline

		\multicolumn{8}{c}{\textbf{Selections}} \\
		\hline 
		U-test (adjusted) & 2.9e-07 & 2.3e-11 & 3.2e-01 & 1.2e-01 & 4.3e-03 & 4.5e-01 & 1.2e-01 \\
		\hline
		Cohen's $d$ & 0.23 & 0.26 & 0.01 & 0.09 & 0.09 & 0.07 & 0.06 \\  
		\hline

		\multicolumn{8}{c}{\textbf{Edit ratio}} \\
		\hline 
		U-test (adjusted) & 3.5e-05 & 3.3e-02 & 3.3e-02 & 2.0e-01 & 4.0e-03 & 1.4e-01 & 5.9e-03 \\
		\hline 
		Cohen's $d$ & 0.31 & 0.13 & 0.10 & 0.06 & 0.17 & 0.09 & 0.19  \\
		\hline

@@

%\begin{table}[ht!]
%	\tiny
%	\renewcommand{\arraystretch}{1.3}
%	\caption{Effect size and statistical significance of the relationships between edits and selection per minute, edit ratio, and duration of interruptions in Mylyn} %, and editions and selections.}
%	\label{tbl:effect_size_dur_mylyn}
%	\centering
%	\begin{tabular}{l | p{1cm} | p{1cm} | p{1cm} } 
%		& low & medium & high  \\  
%		\hline
%		\multicolumn{4}{c}{\textbf{Edits}} \\
%		\hline
%		mean & 1.73 & 1.65 & 1.87 \\ 
%		% \cline{3-4} 
%		%t.test & $\hookrightarrow$& \multicolumn{2}{c}{$<$ 2.2e-16} \\
%		\cline{3-4} 
%		U-test & $\hookrightarrow$ & 3.71e-05 & 0.088  \\
%		
%		\cline{3-4} 
%		Cohen's $d$ & $\hookrightarrow$	& \textbf{0.03} & \textbf{0.06} \\
%		\hline
%		
%		
%		\multicolumn{4}{c}{\textbf{Selections}} \\
%		\hline 
%		mean & 2.21 & 2.18 & 2.53\\ 
%		% \cline{3-4} 
%		%t.test & $\hookrightarrow$& 2.089e-08 & 8.175e-14 \\
%		\cline{3-4} 
%		U-test & $\hookrightarrow$ & 0.000378 & 0.2131 \\
%		
%		\cline{3-4} 
%		Cohen's $d$ & $\hookrightarrow$	& \textbf{0.014} & \textbf{0.14}  \\  
%		\hline
%		\multicolumn{4}{c}{\textbf{Edit ratio}} \\
%		\hline 
%		mean & 0.36 & 0.35 & 0.34 \\ 
%		%   \cline{3-4} 
%		%  t.test & $\hookrightarrow$&  4.20E-12 & $<$ 2.2e-16 \\
%		\cline{3-4} 
%		U-test & $\hookrightarrow$ & 0.2717 & 0.1955 \\
%		\cline{3-4} 
%		Cohen's $d$ & $\hookrightarrow$ & \textbf{0.04} & \textbf{0.07} \\
%		\hline
%		
%	\end{tabular}
%\end{table}

%\begin{table}[ht!]
%	\tiny
%	\renewcommand{\arraystretch}{1.3}
%	\caption{Effect size and statistical significance of the relationships between edits and selection per minute, edit ratio, and duration of interruptions in UDC} %, and editions and selections.}
%	\label{tbl:effect_size_dur_udc}
%	\centering
%	\begin{tabular}{l | p{1cm} | p{1cm} | p{1cm} } 
%		& low & medium & high  \\  
%		\hline
%		\multicolumn{4}{c}{\textbf{Edits}} \\
%		\hline
%		mean & 2.22 & 1.94 & 2.06 \\ 
%		% \cline{3-4} 
%		%t.test & $\hookrightarrow$& \multicolumn{2}{c}{$<$ 2.2e-16} \\
%		\cline{3-4} 
%		U-test & $\hookrightarrow$ & \multicolumn{2}{c}{$<$ 2.2e-16}  \\
%		
%		\cline{3-4} 
%		Cohen's $d$ & $\hookrightarrow$	& \textbf{0.12} & \textbf{0.06} \\
%		\hline
%		
%		
%		\multicolumn{4}{c}{\textbf{Selections}} \\
%		\hline 
%		mean & 2.03 & 1.85 & 1.97\\ 
%		% \cline{3-4} 
%		%t.test & $\hookrightarrow$& 2.089e-08 & 8.175e-14 \\
%		\cline{3-4} 
%		U-test & $\hookrightarrow$ & 3.05e-09 & 1.25e-07 \\
%		
%		\cline{3-4} 
%		Cohen's $d$ & $\hookrightarrow$	& \textbf{0.11} & \textbf{0.03}  \\  
%		\hline
%		\multicolumn{4}{c}{\textbf{Edit ratio}} \\
%		\hline 
%		mean & 0.47 & 0.46 & 0.454 \\ 
%		%   \cline{3-4} 
%		%  t.test & $\hookrightarrow$&  4.20E-12 & $<$ 2.2e-16 \\
%		\cline{3-4} 
%		U-test & $\hookrightarrow$ & 2.57e-07 & 5.11e-14  \\
%		\cline{3-4} 
%		Cohen's $d$ & $\hookrightarrow$ & \textbf{0.11} & \textbf{0.15} \\
%		\hline
%		
%	\end{tabular}
%\end{table}

>>>>

In this section we analyze whether or not the duration of the interruptions is a factor in the relationship between interruptions and developer productivity. Under the hypothesis that prolonged interruptions require more time to recover, we built two groups of sessions based on the proportion of long interruptions. As we described previously we consider that a prolonged interruption has a duration of $\geq$ 12 minutes, according to previous work \cite{GM04, KaptelininN07}. Prolonged interruptions could be due to an extended interruption, such as a coffee break or a meeting.

@@

Following the conclusions on RQ1, the number of interruptions has an impact on the metrics and is a variable that should not be ignored. So, we split every group in accordance to the number of interruptions: this time, instead of taking the quantiles of the distribution to create groups we created a group for every number of interruptions up to seven. Since the results in RQ1 show that as the number of interruptions increases the difference between the groups minimizes, we believe that having more distinctive groups instead of groups created by ranges is also valid. Thus, for the three metrics, we created the groups of sessions according to the number on interruptions, between one and seven, and then we split every group in low and high, according to the proportion of prolonged interruptions.

The Figure \ref{fig:dur_mylyn} shows the median of the metrics when grouping the sessions by the number of interruptions and the proportion of prolonged interruptions with Mylyn, and the Figure \ref{fig:dur_udc} shows the same with UDC. According to our theories, it is better to have a low number of interruptions and a low proportion of prolonged interruptions to have higher rates of productivity, and the results agree with this statement. First, we observe that (as we conclude in RQ1) when the number of interruption increases the median of the metric decreases until reaching a point (approximately between three and four interruptions) where the change is minimal and continues that way for the rest of the groups. Second, almost in every group the median value of groups with low proportion of prolonged interruptions is greater than in groups with a high proportion. The latter observation is more clear in the edits per minute metric in both datasets, for we can see that at some points the medians intertwine in the selections per minutes and edit ratio, specially with Mylyn.

When having a small number of interruptions (between 1 and 3) the difference between sessions with a low or high proportion of prolonged interruptions is more noticeable, and the group of sessions with a low proportion has higher median values. As the number of interruptions increases the differences get smaller, but we observe that the group of low proportion of prolonged interruptions always shows bigger rates. This comments apply on the edits and selections per minute, and to a lesser extent to the edit ratio, where the changes between groups (both by number of interruptions and proportion) are small.

\begin{figure}[!ht]
\centering
\begin{tabular}{c}
\minipage{0.33\textwidth}
\includegraphics[width=\linewidth,clip=, angle=0]{figures/mylyn_lines_dur_emin}  
\endminipage\hfill
\minipage{0.33\textwidth}
\includegraphics[width=\linewidth,clip=, angle=0]{figures/mylyn_lines_dur_smin}  
\endminipage\hfill
\minipage{0.33\textwidth}
\includegraphics[width=\linewidth,clip=, angle=0]{figures/mylyn_lines_dur_eratio} 
\endminipage\hfill
\end{tabular}
\caption{Median values of edits and selection events per minute, and the edit ratio according to the number of interruptions and proportion of prolonged interruptions in Mylyn. }
\label{fig:dur_mylyn}
\end{figure}

\begin{figure}[!ht]
	\centering
	\begin{tabular}{c}
		\minipage{0.33\textwidth}
		\includegraphics[width=\linewidth,clip=, angle=0]{figures/udc_lines_dur_emin}  \\
		\endminipage\hfill
		\minipage{0.33\textwidth}
		\includegraphics[width=\linewidth,clip=, angle=0]{figures/udc_lines_dur_smin}  \\
		\endminipage\hfil
		\minipage{0.33\textwidth}
		\includegraphics[width=\linewidth,clip=, angle=0]{figures/udc_lines_dur_eratio} \\
		\endminipage\hfill
	\end{tabular}
	\caption{Median values of edits and selection events per minute, and the edit ratio according to the number of interruptions and proportion of prolonged interruptions in UDC. }
	\label{fig:dur_udc}
\end{figure}

When comparing the results between the datasets, we can see the same observations described before. Therefore, we can conclude that the duration of the interruption has an impact when having a low number of interruptions, but when the number is bigger the impact is very subtle. We tested the statistical significance of this results and show the results in the Table \ref{tbl:effect_size_dur_mylyn} for Mylyn, and Table \ref{tbl:effect_size_dur_udc} for UDC. For every group according to the number of interruptions, we measured the effect between the low and high groups of sessions, with the Cohen's $d$ test and U-test. The groups with one to three interruptions have moderate size effects, in comparison with the other groups with more interruptions that have lower size effects; this matches our observations via visual inspection on a more defined difference between the low and high groups when having a small number of interruptions. Starting from (approximately) three interruptions, the distributions of the metrics are very alike, except in some cases like the edit ratio with seven interruptions in Mylyn, where the median is almost the same but the distribution of the values are different enough to have a medium effect size. The bold numbers are those with a significant effect size.

\begin{table}[ht!]
	\tiny
	\renewcommand{\arraystretch}{1.3}
	\caption{Effect size and statistical significance of the relationships between edits and selection per minute, edit ratio, and duration of interruptions in Mylyn} %, and editions and selections.}
	\label{tbl:effect_size_dur_mylyn}
	\centering
	\begin{tabular}{l | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm}} 
		Group & 1 & 2 & 3 & 4 & 5 & 6 & 7  \\  
		\hline
		\multicolumn{8}{c}{\textbf{Edits}} \\
		\hline
		U-test (adjusted) & \textbf{5.6e-10} & \textbf{6.5e-07} & \textbf{2.9e-04} & 0.0074 & 0.0021 & 0.32 &  0.39\\
		\hline
		Cohen's $d$ & \textbf{0.62} & 0.38 & 0.20 & 0.08 & 0.21 & 0.17 & 0.45 \\
		\hline

@@

		\multicolumn{8}{c}{\textbf{Edit ratio}} \\
		\hline 
		U-test (adjusted) & \textbf{2.6e-06} & \textbf{5.7e-05} & 0.023 & \textbf{0.64} & 0.023 & 0.49 & 0.15 \\
		\hline 
		Cohen's $d$ & \textbf{0.52} & 0.40 & 0.21 & 0.06 & 0.41 & 0.04 & \textbf{0.56}  \\
		\hline

@@

\begin{table}[ht!]
	\tiny
	\renewcommand{\arraystretch}{1.3}
	\caption{Effect size and statistical significance of the relationships between edits and selection per minute, edit ratio, and duration of interruptions in UDC} %, and editions and selections.}
	\label{tbl:effect_size_dur_udc}
	\centering
	\begin{tabular}{l | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm} | p{1cm}} 
		Group & 1 & 2 & 3 & 4 & 5 & 6 & 7  \\  
		\hline
		\multicolumn{8}{c}{\textbf{Edits}} \\
		\hline
		U-test (adjusted) & \textbf{1.9e-20} & \textbf{4.0e-13} & 0.0011 & 0.027 & \textbf{7.2e-08} & 0.055 & 6.5e-06\\
		\hline
		Cohen's $d$ & 0.28 & 0.25 & 0.09 & 0.03 & 0.11 & 0.01 & 0.15 \\
		\hline

		\multicolumn{8}{c}{\textbf{Selections}} \\
		\hline 
		U-test (adjusted) & \textbf{2.9e-07} & \textbf{2.3e-11} & 0.32 & 0.12 & 0.0043 & 0.45 & 0.12 \\
		\hline
		Cohen's $d$ & 0.23 & 0.26 & 0.01 & 0.09 & 0.09 & 0.07 & 0.06 \\  
		\hline

		\multicolumn{8}{c}{\textbf{Edit ratio}} \\
		\hline 
		U-test (adjusted) & \textbf{3.5e-05} & 0.033 & 0.033 & 0.20 & 0.004 & 0.14 & 0.0059 \\
		\hline 
		Cohen's $d$ & 0.31 & 0.13 & 0.10 & 0.06 & 0.17 & 0.09 & 0.19  \\
		\hline

@@
